Finding Optimal Sequences for Area Aggregation---\protect\\ \Astar vs. Integer Linear Programming

### 2019-03-23 area_agg_seq_tsas.tex
showed exponential lower bound using patches with different sizes

\keywords{Continuous map generalization, Land cover, 
	 Type change, Compactness, Shortest path
%	 , Optimization, Exponential
	 }

@@

We assume that there exist many-to-one relationships 
between the areas of the start map and the areas of the goal map.
This assumption is based on the fact that many algorithms 
\parencite[e.g.,][]{HaunertWolff2010AreaAgg,
vanSmaalen2003,Oehrlein2017Aggregation}
result in many-to-one relationships
when aggregating land-cover areas.
Their inputs and generalized results together
can be used as our inputs.
However, we should not use those algorithms
to generate a sequence of maps at different scales, 
because those algorithms do not take into account 
the consistence of the generated maps.
We term the areas of the goal map \emph{regions}.
That is, every region is the union of a set of areas 
on the start map.
The type of a region may differ from the types of its components. 
For example, a small water area together with 
multiple adjacent forest areas may constitute 
a large forest region on the smaller-scale map.
However, we assume that every region, on the goal map, 
contains at least one area of the same type on the start map.
Our assumptions hold if the goal map has been produced 
with an automatic method for area aggregation, 
for example, by the method of \citet{HaunertWolff2010AreaAgg}.
That method produces a land-cover map at a single smaller scale, 
given a land-cover map at a larger scale.
Although \textcite{HaunertWolff2010AreaAgg}
attain results of high quality, 
they do not produce a sequence land-cover maps.

@@

\mypar{Contribution}
We show related work (\sect\ref{sec:AreaAgg_RelatedWork}).
Then, we formally model our problem, 
analyze the size of our model in a worst-case scenario,
introduce our methods,
and present the basic concepts of our method
(\sect\ref{sec:AreaAgg_Preliminaries}).
We define our cost functions 
(\sect\ref{sec:AreaAgg_CostFunctions}).
We prove that our problem is NP-hard
(\sect\ref{sec:AreaAgg_NP-hardness}).
Then, we develop and compare three methods 
for finding aggregation sequences.
First, we present 
a greedy algorithm (\sect\ref{sec:AreaAgg_Greedy}).
Second, we develop a new global optimization approach
based on the \Astar algorithm 
(\sect\ref{sec:AreaAgg_AStar}).
Third, we model our pathfinding problem as
an \emph{integer linear program} (ILP), 
and we solve this ILP with minimizing our cost function
(\sect\ref{sec:AreaAgg_ILP}).
Our ILP uses binary (0--1) variables. 
These variables help us to model our problem, 
but in general, it is NP-hard 
to solve an ILP optimally.
By comparing with the greedy algorithm, 
which is used as a benchmark,
we are able to see whether it is worth 
to use \Astar or the ILP-based algorithm, 
which are more complex and slower.  
Our case study 
uses a dataset of the German topographic database ATKIS 
(\sect\ref{sec:AreaAgg_CaseStudy}).
In the conclusions, we show possible ways to improve our 
methods (\sect\ref{sec:AreaAgg_Conclusions}).

@@

We show how to compute an aggregation sequence 
for a single region, $R$. 
For a goal map with many regions,
we ``interleave'' the sequences for each of them
with respect to the order of the smallest patches 
(see for example \fig\ref{fig:AreaAgg_IntegrateSequence}).
This integration is similar to the merge step in the 
Mergesort algorithm; 
see \textcite[pp.~29--37]{Cormen2009}.
To allow us to describe our method more easily,
below we assume that the goal map has only one region.
This region consists of~$n$ land-cover 
areas (components) on the start map. 
In other words, the union of the~$n$ land-cover areas 
is the only region on the goal map.

@@

\mypar{Model}
We consider a directed graph $G_\mathrm{S}$, 
which we call the \emph{subdivision graph} 
(see \fig\ref{fig:AreaAgg_SubdivisionName}). 
The node set $V_\mathrm{S}$ of~$G_\mathrm{S}$ 
contains nodes for all the
possible maps (or \emph{subdivisions}), including the start map, 
all possible intermediate-scale maps, and the goal map.
The arc set~$E_\mathrm{S}$ of~$G_\mathrm{S}$ 
contains an arc $(\Pnode, P_{t+1,j})$ 
between any two maps~$\Pnode$ and~$P_{t+1,j}$ in~$V_\mathrm{S}$ 
if~$P_{t+1,j}$ can be reached from~$\Pnode$ 
with a single aggregation operation,
involving a smallest area.
On this basis, any directed path in~$G_\mathrm{S}$ 
from the start map to the goal map
defines a possible aggregation sequence.

@@

\mypar{Exponential lower bound}
We now analyze the size of subdivision graph~$G_\mathrm{S}$.
Our analysis is inspired by \textcite{Keane1975Size},
where we use a row of~$n$ land-cover areas.
In our instance
(see \fig\ref{fig:AreaAgg_ExponentialSize}), 
the start map consists of~$n=2k+1$ rectangles,
and the goal map is simply the union of the~$n$ rectangles.
At layer~$L_{k+1}$, we need to remove~$k$ of the~$2k$ 
intermediate vertical boundaries. 
For this, the number of choices is~$
\binom{2k}{k}
= \frac{2k\cdot (2k-1)\cdot \ldots \cdot (k+1)}
{k\cdot (k-1)\cdot \ldots \cdot 1}
\ge 2^k = 2^{(n-1)/2}$.
That is to say, the size of our model
has exponential lower bound.

\mypar{Methods}
Our idea is to obtain an optimal aggregation sequence through 
computing a minimum-weight path from the start to the goal 
(see \fig\ref{fig:AreaAgg_SubdivisionName}).
This idea obviously requires that the arc weights are 
set such that a minimum-weight start--goal
path does actually correspond to an aggregation sequence of 
maximum cartographic quality.  
Moreover, putting the idea to practice is far from trivial 
since graph~$G_\mathrm{S}$ can be huge.
We compare a greedy algorithm, \Astar, and an ILP-based algorithm
in finding such paths.
Note that our inputs are only
subdivisions~$\Pstart$ and~$\Pgoal$.
We generate a subdivision (node) only when we want to visit it.

In directed acyclic graphs, shortest paths can be found slightly faster 
than in general directed or undirected graphs.
An off-the-shelf shortest-path algorithm for directed acyclic graphs,
however, will explore the whole graph, which has exponential size.
The challenge in our work was to tune the A* algorithm such that it
explores only a small fraction of the graph.

\mypar{Notation}
We represent each land-cover area by a polygon with a type.
We denote by $P$ the set of polygons on the start map.
We use~$p$, $q$, $r$, or~$o$ to denote polygons.
A \emph{patch} is a set of polygons whose union is connected. 
Each patch also has a unique land-cover area type.
We use~$u$ or~$v$ to denote the patches.

@@

\begin{figure}[tb]
\centering
\includegraphics[page=4]{AreaAgg_Preliminaries}
\caption{An example to show that the size of our problem 
	has an exponential lower bound.}
\label{fig:AreaAgg_ExponentialSize}
\end{figure}

@@

>>>>

\keywords{Continuous map generalization, Land cover, 
	 Type change, Compactness, Shortest path
%	 , Optimization, Exponential, NP-hard
	 }

@@

We assume that there exist many-to-one relationships 
between the areas of the start map and the areas of the goal map.
This assumption is based on the fact that many algorithms 
\parencite[e.g.,][]{HaunertWolff2010AreaAgg,
vanSmaalen2003,Oehrlein2017Aggregation}
result in many-to-one relationships
when aggregating land-cover areas.
Their inputs and generalized results together
can be used as our inputs.
However, we should not use those algorithms
to generate a sequence of maps at different scales
because those algorithms do not take into account 
the consistence of the generated maps.
We term the areas of the goal map \emph{regions}.
That is, every region is the union of a set of areas 
on the start map.
The type of a region may differ from the types of its components. 
For example, a small water area together with 
multiple adjacent forest areas may constitute 
a large forest region on the smaller-scale map.
However, we assume that every region, on the goal map, 
contains at least one area of the same type on the start map.
Our assumptions hold if the goal map has been produced 
with an automatic method for area aggregation, 
for example, by the method of \citet{HaunertWolff2010AreaAgg}.
That method produces a land-cover map at a single smaller scale, 
given a land-cover map at a larger scale.
Although \textcite{HaunertWolff2010AreaAgg}
attain results of high quality, 
they do not produce a sequence land-cover maps.

@@

\mypar{Contribution}
We show related work (\sect\ref{sec:AreaAgg_RelatedWork}).
Then, we formally model our problem, 
analyze the size of our model in a worst-case scenario,
introduce our methods,
and present the basic concepts of our method
(\sect\ref{sec:AreaAgg_Preliminaries}).
We define our cost functions 
(\sect\ref{sec:AreaAgg_CostFunctions}).
We prove that our problem is NP-hard
(\sect\ref{sec:AreaAgg_NP-hardness}).
Then, we develop and compare three methods 
for finding aggregation sequences.
First, we present 
a greedy algorithm (\sect\ref{sec:AreaAgg_Greedy}).
Second, we develop a new global optimization approach
based on the \Astar algorithm 
(\sect\ref{sec:AreaAgg_AStar}).
Third, we model our pathfinding problem as
an \emph{integer linear program} (ILP), 
and we solve this ILP with minimizing our cost function
(\sect\ref{sec:AreaAgg_ILP}).
Our ILP uses binary (0--1) variables. 
These variables help us to model our problem, 
but in general, it is NP-hard 
to solve an ILP optimally.
By comparing with the greedy algorithm, 
which is used as a benchmark,
we are able to see whether it is worth 
to use \Astar or the ILP-based algorithm, 
which are more complex and slower.  
Our case study 
uses a dataset of the German topographic database ATKIS 
(\sect\ref{sec:AreaAgg_CaseStudy}).
In the concluding remarks, we show possible ways to improve our 
methods (\sect\ref{sec:AreaAgg_Conclusions}).

@@

We show how to compute an aggregation sequence 
for a single region, $R$. 
For a goal map with many regions,
we ``interleave'' the sequences for each of them
with respect to the order of the smallest patches 
(see for example \fig\ref{fig:AreaAgg_IntegrateSequence}).
This integration is similar to the merge step in the 
Mergesort algorithm; 
see \textcite[\sect2.3]{Cormen2009}.
To allow us to describe our method more easily,
below we assume that the goal map has only one region.
This region consists of~$n$ land-cover 
areas (components) on the start map. 
In other words, the union of the~$n$ land-cover areas 
is the only region on the goal map.

@@

\subsection{Model}
\label{sec:AreaAgg_Model}
We consider a directed graph $G_\mathrm{S}$, 
which we call the \emph{subdivision graph} 
(see \fig\ref{fig:AreaAgg_SubdivisionName}). 
The node set $V_\mathrm{S}$ of~$G_\mathrm{S}$ 
contains nodes for all the
possible maps (or \emph{subdivisions}), including the start map, 
all possible intermediate-scale maps, and the goal map.
The arc set~$E_\mathrm{S}$ of~$G_\mathrm{S}$ 
contains an arc $(\Pnode, P_{t+1,j})$ 
between any two maps~$\Pnode$ and~$P_{t+1,j}$ in~$V_\mathrm{S}$ 
if~$P_{t+1,j}$ can be reached from~$\Pnode$ 
with a single aggregation operation,
involving a smallest area.
On this basis, any directed path in~$G_\mathrm{S}$ 
from the start map to the goal map
defines a possible aggregation sequence.

@@

\subsection{Notation}
\label{sec:AreaAgg_Notation}

We represent each land-cover area by a polygon with a type.
We denote by~$P$ the set of polygons on the start map.
We use~$p$, $q$, $r$, or~$o$ to denote polygons.
A \emph{patch} is a set of polygons whose union is connected. 
Each patch also has a unique land-cover area type.
We use~$u$ or~$v$ to denote the patches.

@@

\begin{figure}[tb]
\centering
\includegraphics[page=4]{AreaAgg_Preliminaries}
\caption{An example to show that 
    the size of subdivision graph~$G_\mathrm{S}$
	has exponential lower bound.}
\label{fig:AreaAgg_ExponentialSize}
\end{figure}

@@

\subsection{Exponential lower bound}
\label{sec:AreaAgg_Exponential}
We now analyze the size of subdivision graph~$G_\mathrm{S}$.
Our analysis is inspired by \textcite{Keane1975Size},
where we use a row of~$n$ land-cover areas.
In our instance
(see \fig\ref{fig:AreaAgg_ExponentialSize}), 
the start map consists of~$n=2k+1$ rectangular patches,
and the goal map is simply the union of the~$n$ patches.
The patches have area 
sizes~$
%\left( 
100 + \frac{1}{n}, 99+ \frac{2}{n}, 
100 + \frac{3}{n}, 99+ \frac{4}{n}, \ldots, 
99 + \frac{n-1}{n}, \text{and~} 101 
%\right)
$, 
from left to right.
According to our rule, we always 
aggregate the smallest patch with one of its neighbors.
Therefore, in the first~$k$ steps from the start map,
we aggregate every other patch with one of its neighbors.
At each of the steps, we have two choices,
aggregating with the left patch or with the right one.
As a result, there are~$2^k = 2^{(n-1)/2}$ subdivisions 
in layer~$L_{k+1}$.
That is to say, the size of subdivision graph~$G_\mathrm{S}$
has exponential lower bound.

\subsection{Methods}
\label{sec:AreaAgg_Methods}

Our idea is to obtain an optimal aggregation sequence through 
computing a minimum-weight path from the start to the goal 
(see \fig\ref{fig:AreaAgg_SubdivisionName}).
This idea obviously requires that the arc weights are 
set such that a minimum-weight start--goal
path does actually correspond to an aggregation sequence of 
maximum cartographic quality.  
Moreover, putting the idea to practice is far from trivial 
since graph~$G_\mathrm{S}$ can be huge.
We compare a greedy algorithm, \Astar, and an ILP-based algorithm
in finding such paths.
Note that our inputs are only
subdivisions~$\Pstart$ and~$\Pgoal$.
We generate a subdivision (node) only when we want to visit it.

In directed acyclic graphs, shortest paths can be found slightly faster 
than in general directed or undirected graphs.
An off-the-shelf shortest-path algorithm for directed acyclic graphs
(e.g., \textcite[\sect25.2]{Cormen2009}),
however, will explore the whole graph, which has exponential size.
The challenge in our work was to tune the A* algorithm such that it
explores only a small fraction of the graph.
